---
title: "Do Homeowners Value Air Quality: Evidence From Wildfire Smoke"
author: "Tristan Misko"
date: "1 December 2021"
output: pdf_document
header-includes:
    - \usepackage{setspace}\doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(ggplot2)
library(data.table)
library(stargazer)
library(biglm)
library(usmap)
```

```{r}
#import data
main <- read.csv("/Users/tristanmisko/Documents/Berkeley/ECON/191/2021/Data/Clean\ Data/main.csv", colClasses = c("fips"="character"))
# function which record treatment status based on the smoke score differential
set.status <- Vectorize(function(value, control.threshold, treatment.threshold){
    if (abs(value) < control.threshold){
        return("control")
    } else if (value > treatment.threshold){
        return("positive treatment")
    } else if (value < -treatment.threshold){
        return("negative treatment")
    } else {
        return("boundary")
    }
}, "value")

# perform the analysis for t* = 55
main$period <- as.factor(main$period)
data.pch43 <- select(main, -c(1:3,25:40))
data.pch48 <- select(main, -c(1:3,21:24,29:40))
data.pch55 <- select(main, -c(1:3,21:28,33:40))
data.pch60 <- select(main, -c(1:3,21:32,37:40))
data.pch67 <- select(main, -c(1:3,21:36))

data.pch55$period <- as.factor(data.pch55$period)

# get status at different threshold values
data.pch43$status.k5.t25 <- as.factor(set.status(data.pch43$m.s.pch43, 0.05, 0.25))
data.pch43$status.k5.t50 <- as.factor(set.status(data.pch43$m.s.pch43, 0.05, 0.50))
data.pch43$status.k10.t25 <- as.factor(set.status(data.pch43$m.s.pch43, 0.10, 0.25))
data.pch43$status.k10.t50 <- as.factor(set.status(data.pch43$m.s.pch43, 0.10, 0.50))

data.pch48$status.k5.t25 <- as.factor(set.status(data.pch48$m.s.pch48, 0.05, 0.25))
data.pch48$status.k5.t50 <- as.factor(set.status(data.pch48$m.s.pch48, 0.05, 0.50))
data.pch48$status.k10.t25 <- as.factor(set.status(data.pch48$m.s.pch48, 0.10, 0.25))
data.pch48$status.k10.t50 <- as.factor(set.status(data.pch48$m.s.pch48, 0.10, 0.50))

data.pch55$status.k5.t25 <- as.factor(set.status(data.pch55$m.s.pch55, 0.05, 0.25))
data.pch55$status.k5.t50 <- as.factor(set.status(data.pch55$m.s.pch55, 0.05, 0.50))
data.pch55$status.k10.t25 <- as.factor(set.status(data.pch55$m.s.pch55, 0.10, 0.25))
data.pch55$status.k10.t50 <- as.factor(set.status(data.pch55$m.s.pch55, 0.10, 0.50))

data.pch60$status.k5.t25 <- as.factor(set.status(data.pch60$m.s.pch60, 0.05, 0.25))
data.pch60$status.k5.t50 <- as.factor(set.status(data.pch60$m.s.pch60, 0.05, 0.50))
data.pch60$status.k10.t25 <- as.factor(set.status(data.pch60$m.s.pch60, 0.10, 0.25))
data.pch60$status.k10.t50 <- as.factor(set.status(data.pch60$m.s.pch60, 0.10, 0.50))

data.pch67$status.k5.t25 <- as.factor(set.status(data.pch67$m.s.pch67, 0.05, 0.25))
data.pch67$status.k5.t50 <- as.factor(set.status(data.pch67$m.s.pch67, 0.05, 0.50))
data.pch67$status.k10.t25 <- as.factor(set.status(data.pch67$m.s.pch67, 0.10, 0.25))
data.pch67$status.k10.t50 <- as.factor(set.status(data.pch67$m.s.pch67, 0.10, 0.50))
```


```{r}
sens.data.pch60 <- data.frame(upper = seq(0.10,0.75, by = 0.05))
coef.vector <- c()
thresholds.sens.pch60 <- data.frame(data.pch60)

# 
for (u in sens.data.pch60$upper){
    print(u)
    df <- cbind(thresholds.sens.pch60,
                s = as.factor(set.status(thresholds.sens.pch60$m.s.pch60, 0.05, u)))
    df <- filter(df, (s == "positive treatment") | (s == "control"))
    df$treatment <- as.numeric(df$s == "positive treatment")*
        as.numeric(df$post60)
    reg.out <- biglm(data = df, 
                     zhvi.score ~ treatment + unemp + period + fips)
    coef.vector <- c(coef.vector, coef(reg.out)[2])
}
sens.data.pch60$coef <- coef.vector
coef.vector

```

```{r}
ggplot(data = sens.data.pch60, aes(x = upper, y = coef)) + geom_line() +
    labs(title = "Sensitivity of Changes in Upper Threshold", 
         subtitle = "The Estimate Reflects Significant Thresholding Behavior")
```




```{r, fig.width = 7, fig.height = 5}
# tabulate and plot status by county
plt.pch55.status <- data.pch55 %>% group_by(fips) %>% 
                        summarize(status.k5.t25 = first(status.k5.t25),
                                  status.k5.t50 = first(status.k5.t50),
                                  status.k10.t25 = first(status.k10.t25),
                                  status.k10.t50 = first(status.k10.t50))
#plot map for kappa = 5, tau = 50
plot_usmap("counties", data = plt.pch55.status, values = "status.k5.t50",
           include = unique(plt.pch55.status$fips)) + 
    labs(title = "Treatment Status by County",
         subtitle = "Control Threshold = 5%, Treatment Threshold = 50%",
         fill = "Treatment Status") +
    theme(legend.position = "right")
```

# Methodology

```{r}
# regression!!!!
#reg.pch55.ols <- biglm(data = data.pch55,
                       #zhvi.score ~ n.score + unemp + period + fips)

#data.pch55$treat.k5.t25 <- as.numeric(data.pch55$post55 & 
                           #(data.pch55$status.k5.t25 == "positive treatment"))
#data.pch55$treat.k5.t50 <- as.numeric(data.pch55$post55 & 
#                           (data.pch55$status.k5.t50 == "positive treatment"))
#data.pch55$treat.k10.t25 <- as.numeric(data.pch55$post55 & 
#                            (data.pch55$status.k10.t25 == "positive treatment"))
#data.pch55$treat.k10.t50 <- as.numeric(data.pch55$post55 & 
                            #(data.pch55$status.k10.t50 == "positive treatment"))

#data.pch55.k5.t25 <- filter(data.pch55,
                            #status.k5.t25 %in% c("positive treatment", "control")) 
#reg.pch55.k5.t25 <- biglm(data = data.pch55.k5.t25, 
                          #zhvi.score ~ treat.k5.t25 + unemp + period + fips)
data.pch55.k5.t50 <- filter(data.pch55, 
                            status.k5.t50 %in% c("positive treatment", "control")) 
#reg.pch55.k5.t50 <- biglm(data = data.pch55.k5.t50, 
                          #zhvi.score ~ treat.k5.t50 + unemp + period + fips)
#data.pch55.k10.t25 <- filter(data.pch55, 
                            #status.k10.t25 %in% c("positive treatment", "control")) 
#reg.pch55.k10.t25 <- biglm(data = data.pch55.k10.t25, 
                          #zhvi.score ~ treat.k10.t25 + unemp + period + fips)
#data.pch55.k10.t50 <- filter(data.pch55, 
                            #status.k10.t50 %in% c("positive treatment", "control")) 
#reg.pch55.k10.t50 <- biglm(data = data.pch55.k10.t50, 
                          #zhvi.score ~ treat.k10.t50 + period + unemp + fips)
```

# Results

We summarize the results of the two regressions above in Table 1.  

\begin{table}[!htbp] \centering 
  \caption{Effect of Wildfire Smoke on Housing Prices} 
  \label{} 
\begin{tabular}{@{\extracolsep{6pt}}lcc} 
\\\hline  \hline \\ 
\\ & (1) & (2)\\ 
\hline \\[-1.8ex] 
 (Intercept) & 183687.00$^{***}$ & 73976.901 \\ 
  & (1100.8998) & (1113.0914) \\ 
  & & \\ 
 n.score & $-$0.6631 &  \\ 
  & (0.6412) &   \\ 
  & &  \\ 
 treat.k5.t50 & & $-$1484.0340$^{***}$  \\ 
  & & (168.9167) \\ 
  & & \\ 
 unemp & 1114.7302$^{***}$ & 1563.9206$^{***}$  \\ 
  & (27.7190) & (48.1914)  \\ 
  & & \\ 
 County Fixed Effects& Yes & Yes  \\ 
  &  &    \\ 
 Time Fixed Effects & Yes & Yes \\ 
  &  &    \\ 
 
\hline \\[-1.8ex] 
Observations & 234600 & 45655 \\ 
R$^{2}$ & 0.9804 & 0.98367 \\ 
\hline 
\hline \\[-1.8ex]  \\ 
\end{tabular} 
\end{table} 

The coefficient from equation (1), the standard OLS equation, is not significantly
different from zero.  Its interpretation is that a one point increase in the smoke
score in a given month, roughly corresponding to a 1 $\mu g$ increase in wildfire
smoke exposure over the course of the month, is associated with a $-0.6412$ drop
in ZHVI.

The coefficient on equation (2) is negative and highly significant.  Its 
interpretation is that the ZHVI (roughly interpreted as the smoothed median
home price in the county in dollars) is expected to fall by $1484.0340$ given
a 50\% long term increase in wildfire smoke exposure.  Here, long term refers
to the difference between the 2010-2014 and the 2015-2019 means.


## Robustness Checks 

Robustness checks will be a major feature of my paper because of the number of
threshold choices I have had to make in the course of my analysis.

\textbf{Parallel Trends:} We plot the mean ZHVI across the treatment and control 
groups in each period and examine the evolution in time.  From the graph, it is
visually clear that the trends in the means are close to parallel, and it seems
that there is a small dip in the ZHVI for the treatment group after treatment. 

```{r}
# parallel trends plot 
plt.means <- data.pch55.k5.t50 %>% group_by(status.k5.t50, period) %>% 
    summarize(m.z = mean(zhvi.score))
plt.means$period <- as.numeric(plt.means$period)

parallel <- ggplot(NULL, aes(x = period, y = m.z)) + 
    geom_vline(aes(xintercept = 55), linetype = "dashed") +
    geom_line(data = filter(plt.means, status.k5.t50 == "positive treatment"),
              aes(col = "treated")) +
    geom_line(data = filter(plt.means, status.k5.t50 == "control"),
              aes(col = "control")) +
    xlab("Period") + ylab("Mean ZHVI") + 
    labs(title = "Evolution of Mean ZHVI by Treatment Group", 
         subtitle ="Trends Before Treatment")
parallel  
```

\textbf{Effect Size Threshold for Treatment Status:} One major choice which will
likely have significant bearing on the resultant regression coefficient is the treatment status.  Changing this threshold affects which counties are included in the 
treatment and control groups.  To address robustness, I plan to produce a plot
which varies the threshold continuously on the $x$ axis and plot the resultant regression coefficient on the $y$ axis.  Tbis plot will give visual insight 
into the sensitivity of the effect size estimate to the threshold choice.

\textbf{Treatment Start Time:} I will also run the models with different periods
in which the treatment begins.





## 

The geographic clustering of treatment effects represents the most significant 
threat to internal validity.  Because smoke density is a roughly continuous
variable, the treatment status of neighboring counties is a strong predictor of
the .  This threat is particularly salient for the simple treatment/control
group model.  